---
title: "Getting Started with ethsign"
author: "Dereck Mezquita"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ethsign}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

# Signing is OFFLINE and deterministic (RFC 6979): this vignette executes with
# no network, no funds, and no chain connection. Every signature printed below
# is the real, reproducible output of the code shown.
box::use(ethsign[eth_signer, eth_address, keccak256, eip712_digest, ecrecover, as_rsv, as_hex])
```

`ethsign` provides the cryptographic primitives needed to sign Ethereum and EVM
messages in pure R: keccak-256 hashing, secp256k1 ECDSA (with the Ethereum
recovery id and EIP-2 low-s normalisation), EIP-712 typed-data digests, EIP-191
`personal_sign`, and address derivation.

It is **not** a wallet: there are no funds, no balances, no chain connection,
and no broadcasting. It is the signing maths plus a small [EthSigner] object
that holds a private key in memory. What you do with the resulting signature is
up to your other tooling.

> **A note on responsibility.** This package handles private keys and produces
> signatures that can authorize real transactions. You are responsible for how
> you use it and for keeping your keys safe.

## Installation

```{r install, eval = FALSE}
renv::install("dereckscompany/ethsign")

# or, if you use remotes instead of renv:
# install.packages("remotes")
# remotes::install_github("dereckscompany/ethsign")
```

## Creating a signer

An `EthSigner` holds a secp256k1 private key privately and signs on its behalf.
The key is normalised to `raw(32)` at construction, never exposed by any public
member, and never printed.

In real use, read the key from the `ETH_PRIVATE_KEY` environment variable so it
never appears in code or history:

```{r signer-env, eval = FALSE}
signer <- eth_signer() # reads ETH_PRIVATE_KEY
```

For a throwaway key (tests, scratch work), use `eth_signer_random()`. For this
vignette we pass an explicit, well-known test key so the printed output is
reproducible -- this is the canonical eth_account / Hyperliquid example key:

```{r signer}
testkey <- "0x0123456789012345678901234567890123456789012345678901234567890123"
signer <- eth_signer(testkey)

signer$address
signer
```

The `$address` is the lowercase `0x`-prefixed last 20 bytes of
`keccak256(pubkey_x || pubkey_y)`. You can derive it without a signer too:

```{r address}
eth_address(testkey)
```

## Hashing

Ethereum uses the original (pre-FIPS-202) Keccak, not SHA3-256. `keccak256()`
accepts a UTF-8 string or raw bytes and returns a plain `raw(32)`:

```{r keccak}
keccak256("") # the canonical empty-string Ethereum vector
keccak256(as.raw(c(0x12, 0x34)))
```

## Signing EIP-712 typed data

EIP-712 is the dominant signing scheme on EVM venues. `ethsign` builds the
`0x1901`-prefixed signing digest for a single struct over atomic field types
(`string`, `uint64`, `uint256`, `bool`, `address`, `bytes32`).

The domain is a named list; the types are an **ordered** list of
`list(name, type)`; the message supplies the field values by name. Here is a
Hyperliquid `usdSend` user action, an exact reproduction of the official
Hyperliquid Python SDK testnet vector:

```{r typed-data}
domain <- list(
  name = "HyperliquidSignTransaction",
  version = "1",
  chainId = 421614,
  verifyingContract = "0x0000000000000000000000000000000000000000"
)

types <- list(
  list(name = "hyperliquidChain", type = "string"),
  list(name = "destination", type = "string"),
  list(name = "amount", type = "string"),
  list(name = "time", type = "uint64")
)

message <- list(
  hyperliquidChain = "Testnet",
  destination = "0x5e9ee1089755c3435139848e47e6635505d5a13a",
  amount = "1",
  time = 1687816341423
)

sig <- signer$sign_typed_data(domain, "HyperliquidTransaction:UsdSend", types, message)
sig
```

If you need the digest on its own (for example to verify it, or to sign it
later with `$sign_digest()`), compute it directly with `eip712_digest()`:

```{r digest}
digest <- eip712_digest(domain, "HyperliquidTransaction:UsdSend", types, message)
digest
```

## The two serializations

A signature is the same object regardless of venue; only the wire format
differs. `as_rsv()` gives the `{r, s, v}` object form (e.g. Hyperliquid, with
`r`/`s` as minimal `0x`-hex); `as_hex()` gives the 65-byte concatenated
`r`/`s`/`v` hex string (e.g. Polymarket):

```{r serialize}
as_rsv(sig)

as_hex(sig)
```

## Verifying with `ecrecover`

`ecrecover()` is Ethereum's signature-to-address recovery: given the signed
digest and the `(r, s, v)` triple, it returns the address that produced the
signature. Recovering the test key's own address closes the loop:

```{r recover}
rsv <- as_rsv(sig)
recovered <- ecrecover(digest, rsv$r, rsv$s, rsv$v)

recovered
recovered == signer$address
```

## Signing a login message

`$sign_message()` applies the EIP-191 `personal_sign` prefix
(`"\x19Ethereum Signed Message:\n" + nbytes + text`) before hashing and signing.
This is the digest behind Sign-In with Ethereum (SIWE) and most "sign this
message to log in" flows:

```{r login}
login <- "example.com wants you to sign in with your Ethereum account"
msig <- signer$sign_message(login)

as_hex(msig)
```

A relying party verifies the login the same way -- reconstruct the prefixed
digest and `ecrecover` the claimed address:

```{r login-verify}
body <- charToRaw(enc2utf8(login))
prefix <- c(
  as.raw(0x19),
  charToRaw("Ethereum Signed Message:\n"),
  charToRaw(as.character(length(body)))
)
login_digest <- keccak256(c(prefix, body))

mrsv <- as_rsv(msig)
ecrecover(login_digest, mrsv$r, mrsv$s, mrsv$v) == signer$address
```

## Where to go next

- `?EthSigner` for the full signer API (`$sign_digest`, `$sign_typed_data`,
  `$sign_message`).
- `?eip712_digest` for the typed-data digest details and field-type scope.
- `?ecrecover` for signature verification.

The same primitives cover Hyperliquid user actions, Polymarket orders, 0x / CoW
/ 1inch / Seaport, ERC-2612 and Permit2 permits, Gnosis Safe, Sign-In with
Ethereum, and raw EVM transaction signing.
